---
title: "NHATS_baseline_preprocessing"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(readstata13)
round_1_sp <- read.dta13('/Users/ChristianHaudenschild/Documents/NHATS Analysis/data/NHATS_Round_1_SP_File.dta')
```

```{r}
dim(round_1_sp)
```

##Cohort derivation

###Exclusion Criteria

####1. Baseline dementia 
    - (hc1disescn9) - 1 - YES, 2 - NO, -1 Inapplicable, -8 DK, -9 Missing
    - filter hc1disescn9 == 2
    - n post filter = 7146

```{r}
round_1_cohort = round_1_sp %>%
  filter(hc1disescn9 ==2)
```

####2. Missing grip strength data
    - remove rows missing grip strength data
    - n post filter = 5969
```{r}
#examination of grip strength related fields
# grep(pattern = 'gr1handtstd1',x = colnames(round_1_cohort))
# grep(pattern = 'gr1grp2noat9',x = colnames(round_1_cohort))
# round_1_cohort[,747:764]
```

```{r}
#check if a respondent has reading 2 but not reading 1 - a single respondent 
#soln. moved their reading 2 into their reading 1 field
sum(is.na(round_1_cohort$gr1grp1rdng) & !is.na(round_1_cohort$gr1grp2rdng))

round_1_cohort[is.na(round_1_cohort$gr1grp1rdng) & !is.na(round_1_cohort$gr1grp2rdng),750] = 
round_1_cohort[is.na(round_1_cohort$gr1grp1rdng) & !is.na(round_1_cohort$gr1grp2rdng),759]

#now can safely remove all NAs based on first reading, n = 5969
round_1_cohort = round_1_cohort[complete.cases(round_1_cohort$gr1grp1rdng),]

#check
sum(is.na(round_1_cohort$gr1grp1rdng))
```

####3. Missing BMI data
    - weight in pounds (hw1currweigh), height in feet (hw1howtallft), height in inches (hw1howtallin)
    - filter out missing rows for either field
    - n post filter = 5822

```{r}
round_1_cohort = round_1_cohort[complete.cases(round_1_cohort$hw1currweigh),]
round_1_cohort = round_1_cohort[complete.cases(round_1_cohort$hw1howtallin),]

#check
sum(is.na(round_1_cohort$hw1currweigh))
sum(is.na(round_1_cohort$hw1howtallin))
sum(is.na(round_1_cohort$hw1howtallft))

```

##Derived measures

####1. Max grip strength
    - maximum of two grip strength measurements (gr1grp1rdng and gr1grp2rdng)
    - appended as max_grip
```{r}
round_1_cohort = round_1_cohort %>%
  mutate(max_grip = case_when(gr1grp1rdng >= gr1grp2rdng~gr1grp1rdng,
                              gr1grp1rdng < gr1grp2rdng~gr1grp2rdng,
                              is.na(gr1grp2rdng) ~ gr1grp1rdng))
```

####2. BMI 
    - defined as self-reported baseline weight in kg divided by height in meters-squared
    - appended as weight_kg, height_m, and BMI respectively
```{r}
#append height and weight conversions
round_1_cohort = round_1_cohort %>%
  mutate(weight_kg = hw1currweigh/2.2046, height_m = (hw1howtallft*12 +hw1howtallin)*0.0254) %>%
  mutate(BMI = weight_kg/height_m^2)
  #dplyr::select(hw1currweigh,hw1howtallft,hw1howtallin,weight_kg,height_m,BMI)
```

####3. High waist circumference
    - waist measure in inches (wc1wstmsrinc)
      - convert to cm
    - binary indicator for high waist circumference,  >= 102 cm in males, >= 88 cm in females
    - appended as high_wc
    
```{r}
#explicit w/ case_when statement because of NAs
round_1_cohort = round_1_cohort %>%
  mutate(high_wc = case_when(wc1wstmsrinc*2.54>=102 & r1dgender ==1 ~ 1,
                             wc1wstmsrinc*2.54>=88 & r1dgender ==2 ~ 1,
                             wc1wstmsrinc*2.54<102 & r1dgender ==1 ~ 0,
                             wc1wstmsrinc*2.54<88 & r1dgender ==2 ~ 0))
                             
#round_1_cohort %>%
  #dplyr::select(wc1wstmsrinc,r1dgender,high_wc)

#104 with no waist circumference data
sum(is.na(round_1_cohort$high_wc))
```

####4. Sarcopenia
    - grip strength < 35.5 kg in males, <20 kg in females 
    - appended as sarcopenia

```{r}
round_1_cohort = round_1_cohort %>%
  mutate(sarcopenia = case_when(max_grip < 35.5 & r1dgender == 1 ~ 1,
                                max_grip < 20 & r1dgender == 2 ~ 1,
                                max_grip >= 35.5 & r1dgender == 1 ~ 0,
                                max_grip >= 20 & r1dgender == 2 ~ 0))

#should be zero NAs, as those with missing grip strength info were already excluded
sum(is.na(round_1_cohort$sarcopenia))
```

####5. SDOC Sarcopenia
    - grip strength/BMI < 1.05 in males, < 0.79 in females
    - appended as sdoc_sarcopenia
    
```{r}
round_1_cohort = round_1_cohort %>%
  mutate(sdoc_sarcopenia = case_when((max_grip/BMI) < 1.05 & r1dgender == 1 ~ 1,
                                     (max_grip/BMI) < 0.79 & r1dgender == 2 ~ 1,
                                     (max_grip/BMI) >= 1.05 & r1dgender == 1 ~ 0,
                                     (max_grip/BMI) >= 0.79 & r1dgender == 2 ~ 0))

#again, should be no NAs
sum(is.na(round_1_cohort$sdoc_sarcopenia))
```

####6. Race (sensitive information)
    - White: rl1yourrace1 == 1
    - Black: rl1yourrace2 == 1
    - Asian: rl1yourrace5 == 1
    - Other: rl1yourrace3 == 1 (American Indian)
             rl1yourrace4 == 1 (Alaskan Native)
             rl1yourrace6 == 1 (Native Hawaiian)
             rl1yourrace7 == 1 (Pacific Islander)
             rl1yourrace8 == 1 (Other)
    
    Hispanic race:
    - r1dracehisp
             
```{r}
# round_1_cohort = round_1_cohort %>%
#   mutate(race = case_when(rl1yourrace1 == 1 ~ 'White',
#                           rl1yourrace2 == 1 ~ 'Black',
#                           rl1yourrace5 == 1 ~ 'Asian',
#                           rl1yourrace3 == 1 | rl1yourrace4 == 1 |
#                           rl1yourrace6 == 1 | r11yourrace7 == 1 |
#                           rl1yourrace8 == 1 ~ 'Other'))

round_1_cohort = round_1_cohort %>%
  mutate(race = case_when(rl1dracehisp == 1 ~ 'White',
                          rl1dracehisp == 2 ~ 'Black',
                          rl1dracehisp == 3 ~ 'Other',
                          rl1dracehisp == 4 ~ 'Hispanic'))
```

####7. Smoking status
    - Current - sd1smokedreg == 1 (smoked regularly) & sd1smokesnow == 1 (smokes now)
    - Former smoker - sd1smokedreg == 1 & sd1smokesnow == 2
    - Never - sd1smokedreg == 2 & sd1smokesnow == 2
    - appended as smoking_status
    
```{r}
round_1_cohort = round_1_cohort %>%
  mutate(smoking_status = case_when(sd1smokedreg == 1 & sd1smokesnow == 1 ~ 'Current',
                                    sd1smokedreg == 1 & sd1smokesnow == 2 ~ 'Former',
                                    sd1smokedreg == 2 ~ 'Never'))

#check 
sum(na.omit(round_1_cohort$sd1smokedreg == 2 & round_1_cohort$sd1smokesnow ==1))

#single NA
which(is.na(round_1_cohort$smoking_status))

round_1_cohort[4548,] %>%
  dplyr::select(sd1smokedreg,sd1smokesnow,smoking_status)
```

####8. Education 
    - el1higstschl (note, in prospectus, this is listed as op1leveledu -w
        which is the field corresponding to other person's education - check this)
    - Less than high school: 1 - no schooling,
                             2 - 1st to 8th grade,
                             3 - 9th to 12th grade, no diploma
    - High school to some college: 4 - high school graduate (diploma or equivalent)
                                   5 - vocational, technical, business or trade school certificate 
                                       beyond high school
                                   6 - some college but no degree
    - College degree: 7 - associate's degree
                      8 - bachelor-s degree
    - Graduate degree: 9 - master's, professional, or doctoral
    - appended as education
                    
```{r}
round_1_cohort = round_1_cohort %>%
  mutate(education = case_when(el1higstschl == 1 | el1higstschl == 2|
                               el1higstschl == 3 ~ 'Less than high school',
                               el1higstschl == 4 | el1higstschl == 5|
                               el1higstschl == 6 ~ 'High school to some college',
                               el1higstschl == 7 | el1higstschl == 8 ~
                                 'College degree',
                               el1higstschl == 9 ~ 'Graduate degree'))


#4 NAs

round_1_cohort[is.na(round_1_cohort$el1higstschl),'el1higstschl']
sum(is.na(round_1_cohort$education))                     
```

####Physical activity

- Participation in activities:
  - pa1evrgowalk (1 - YES, 2 - NO)
  - pa1vigoactv (1 - YES, 2 - NO)
  - pa1favact (712 - walking/jogging,
               713 - other outdoor recreating activities,
               799 - other active leisure)
               
```{r}
round_1_cohort = round_1_cohort %>%
  mutate(ever_walk = case_when(pa1evrgowalk == 1 ~ 1,
                               pa1evrgowalk == 2 ~ 0),
         vigor_actv = case_when(pa1vigoractv == 1 ~ 1,
                                pa1vigoractv == 2 ~ 0))
```


####Comorbidities
- heart_disease:
      - hc1disescn1 - had heart attack, 1 - YES 2 - NO or hc1disescn2 - has heart disease, 1 - YES 2 - NO
- hypertension:
      - hc1disescn3 - has high blood pressure, 1 - YES 2 - NO
- arthritis:
      - hc1disescn4 - has arthritis, 1 - YES 2 - NO
- diabetes:
      - hc1disescn6 - has diabetes, 1 - YES 2 - NO
- lung_disease:
      - hc1disescn7 - has lung disease, 1 - YES 2 - NO
- stroke:
      - hc1disescn8 - has had stroke, 1 - YES 2 - NO
- cancer:
      - hc1disescn10 - has cancer, 1 - YES 2 - NO
- appended as written

```{r}
round_1_cohort = round_1_cohort %>%
  mutate(heart_disease = case_when(hc1disescn1 == 1 | hc1disescn2 == 1 ~ 1,
                                   hc1disescn1 == 2 | hc1disescn2 == 2 ~ 0))

#0 NAs in heart disease
#round_1_cohort[is.na(round_1_cohort$hc1disescn1) & is.na(round_1_cohort$hc1disescn2),]
sum(is.na(round_1_cohort$heart_disease))

round_1_cohort = round_1_cohort %>%
  mutate(hypertension = case_when(hc1disescn3 == 1 ~ 1,
                                  hc1disescn3 == 2 ~0))

#7 NAs
sum(is.na(round_1_cohort$hypertension))

round_1_cohort = round_1_cohort %>%
  mutate(arthritis = case_when(hc1disescn4 == 1 ~ 1,
                                  hc1disescn4 == 2 ~0))

#12 NAs
sum(is.na(round_1_cohort$arthritis))

round_1_cohort = round_1_cohort %>%
  mutate(diabetes = case_when(hc1disescn6 == 1 ~ 1,
                                  hc1disescn6 == 2 ~0))

#2 NAs
sum(is.na(round_1_cohort$diabetes))

round_1_cohort = round_1_cohort %>%
  mutate(lung_disease = case_when(hc1disescn7 == 1 ~ 1,
                                  hc1disescn7 == 2 ~0))

#4 NAs
sum(is.na(round_1_cohort$lung_disease))

round_1_cohort = round_1_cohort %>%
  mutate(stroke = case_when(hc1disescn8 == 1 ~ 1,
                                  hc1disescn8 == 2 ~0))

#5 NAs
sum(is.na(round_1_cohort$stroke))

round_1_cohort = round_1_cohort %>%
  mutate(cancer = case_when(hc1disescn10 == 1 ~ 1,
                                  hc1disescn10 == 2 ~0))

#2 NAs
sum(is.na(round_1_cohort$cancer))
```

#### Age category
- r1d2intvrage 
- 65-69 - 1
- 70-74 - 2
- 75-79 - 3
- 80-84 - 4
- 85+ - 5,6
- appended as age_category
```{r}
round_1_cohort = round_1_cohort %>%
  mutate(age_category  = case_when(r1d2intvrage == 1 ~ '65-69',
                                   r1d2intvrage == 2 ~ '70-74',
                                   r1d2intvrage == 3 ~ '75-79',
                                   r1d2intvrage == 4 ~ '80-84',
                                   r1d2intvrage == 5 | r1d2intvrage == 6 ~ '85+'))

#check
sum(is.na(round_1_cohort$age_category))
```

#### Obesity defined by BMI
- BMI >= 30 kg/m^2
- appended as obesity

```{r}
round_1_cohort = round_1_cohort %>%
  mutate(obesity = case_when(BMI>=30~ 1,
                             BMI<30 ~0))

#check
sum(is.na(round_1_cohort$obesity))
```

### AD8 Dementia classification (NHATS definition)

```{r}

ad8_processing = function(sp, ROUND) {
  for (i in 1:nrow(sp)) {
    
    ###inital codes for dementia_class
    #nursing home resident
    if (sp[i,'r1dresid'] == 7) {
      sp[i,'dementia_class'] = -9
    }                             
  
    if (sp[i,'r1dresid'] %in% c(6,8)) {
      sp[i,'dementia_class'] = -1
    }
    #default to NA
    else {
      sp[i,'dementia_class'] = NA
    }
    
    #default AD8 score to NA
    sp[i,'ad8_score'] = NA
    
    #Get AD8 score if respondent type is proxy and dementia class is NA
    if (sp[i,paste('is',ROUND,'resptype',sep = '')] == 2 & is.na(sp[i,'dementia_class'])) {
      
      sp[i,'ad8_score'] = 0
      
      for (j in 1:8) {
        
        item = paste('cp',ROUND,'chgthink',j,sep = '')
        recoded_item = paste('ad8_item_',j,sep = '')
        sp[i,recoded_item] = NA
        
        #Change reported by proxy (response 1) or dementia/alzheimer's reported by proxy (response 3)
        if (sp[i,item] %in% c(1,3)) {
          sp[i,recoded_item] = 1
        }
        
        #No change reported by proxy
        if (!is.na(sp[i,item]) & sp[i,item] == 2) {
          sp[i,recoded_item] = 0
        }
        
        sp[i,'ad8_score'] = sp[i,'ad8_score'] + sp[i,recoded_item]
      }
      
    }
    
    #fill inapplicable AD8 with -1 and AD8 items with NA
    else {
      for (j in 1:8) {
        recoded_item = paste('ad8_item_',j,sep = '')
        sp[i,recoded_item] = -1
        sp[i,paste('ad8_item_',j,sep = '')] = NA
      }
    }
    
    #if AD8 criterion met in previous round, then set AD8 score = 8 - not applicable for round 1
    if (ROUND != 1) {
      if (sp[i,paste('cp',ROUND,'dad8dem',sep = '')] == 1 & is.na(sp[i,'dementia_class'])) {
        sp[i, 'ad8_score'] = 8
      }
    }
    
    #create ad8_binary (0 - does not meet AD8 criterion, 1 - does meet AD8 criterion)
    if (sp[i,'ad8_score'] >= 2 & !is.na(sp[i,'ad8_score'])) {
      sp[i,'ad8_binary'] = 1
    }
    
    else if (sp[i, 'ad8_score'] %in% c(0,1) | 
             (all(is.na(subset(sp[i,],select = ad8_item_1:ad8_item_8))) & is.na(sp[i,'dementia_class']))) {
      
      sp[i,'ad8_binary'] = 0
    }
    
  }
  return(sp)
} 

round_1_cohort = ad8_processing(round_1_cohort,1)
```

### Orientation domain (NHATS definition)

Sum of date recall and president naming

```{r}

orientation_domain_processing = function(sp, ROUND) {
  
  for (i in 1:nrow(sp)) {
  
    dat_string = paste('cg',ROUND,'todaydat', sep = '')
    
    if (ROUND != 4) {
      for (j in 1:4) {
        
        dat_str = paste(dat_string,j,sep = '')
        item_str = paste('date_item_',j, sep = '')
        
        if (sp[i,dat_str] %in% c(-7,2)) {
          sp[i,item_str] = 0 
        }
        
        else if (sp[i, dat_str] %in% c(-1,-9) | is.na(sp[i,dat_str])) {
          sp[i,item_str] = NA
        }
        
        else {
          sp[i,item_str] = 1
        }
      }
    }
    
    else {
      for (j in c(1,2,3,5)) {
        
        dat_str = paste(dat_string,j,sep = '')
        item_str = paste('date_item_',j, sep = '')
        
        if (j == 5) {
          item_str = 'date_item_4'
        }
        
        if (sp[i,dat_str] %in% c(-7,2)) {
          sp[i,item_str] = 0 
        }
        
        else if (sp[i, dat_str] %in% c(-1,-9) | is.na(sp[i,dat_str])) {
          sp[i,item_str] = NA
        }
        
        else  {
          sp[i,item_str] = 1
        }
      }
    }
    
    sp[i,'date_score'] = sp[i,'date_item_1'] + sp[i,'date_item_2'] + sp[i,'date_item_3'] + sp[i,'date_item_4']
    
    #all NAs in date_items and can't speak to SP (cg*speaktosp == 2) result in date_score == NA
    speaktosp = paste('cg',ROUND,'speaktosp',sep = '')
    if (is.na(sp[i,'date_item_1']) & is.na(sp[i,'date_item_2']) & is.na(sp[i,'date_item_3']) & is.na(sp[i,'date_item_4'] & !is.na(sp[i,speaktosp]) &sp[i,speaktosp] == 2)) {
      sp[i,'date_score'] = NA
    }
    
    pres_strings = c(paste('cg',ROUND,'presidna1',sep = ''),
                     paste('cg',ROUND,'presidna3',sep = ''),
                     paste('cg',ROUND,'vpname1',sep = ''),
                     paste('cg',ROUND,'vpname3',sep = ''))
    
    for (j in 1:4) {
      dat_str = pres_strings[j]
      item_str = paste('o_item_',j, sep = '')
      
      if (sp[i,dat_str] == 0 & !is.na(sp[i,dat_str])) {
        sp[i,item_str] = 0
      }
      
      else if (is.na(sp[i,dat_str])) {
        sp[i,item_str] = NA
      }
      
      else if (sp[i,dat_str] == 1) {
        sp[i,item_str] = 1
      }
    }
    
    sp[i,'orientation_score'] = sp[i,'o_item_1'] + sp[i,'o_item_2'] + sp[i,'o_item_3'] + sp[i,'o_item_4']
    
    #all NAs in o_items and can't speak to SP (cg*speaktosp == 2) result in orientation_score == NA
    if (is.na(sp[i,'o_item_1']) & is.na(sp[i,'o_item_2']) & is.na(sp[i,'o_item_3']) & is.na(sp[i,'o_item_4']) & !is.na(sp[i,speaktosp]) &sp[i,speaktosp] == 2) {
      sp[i,'orientation_score'] = NA
    }
    
    sp[i,'orientation_domain'] = sp[i,'date_score'] + sp[i,'orientation_score']
  }
  return(sp)
}


round_1_cohort = orientation_domain_processing(round_1_cohort, 1)
```

#### Clock score preprocessing

** 8) EXECUTIVE FUNCTION DOMAIN: CLOCK DRAWING SCORE **;
r2clock_scorer=cg2dclkdraw;
if cg2dclkdraw in (-2,-9) then r2clock_scorer=.;
if cg2dclkdraw in (-3,-4,-7) then r2clock_scorer=0;
** IMPUTE MEAN SCORE TO PERSONS MISSING A CLOCK **;
if cg2dclkdraw=-9 and cg2speaktosp=1 then r2clock_scorer=2; ** IF PROXY SAID CAN ASK SP **;
if cg2dclkdraw=-9 and cg2speaktosp=-1 then r2clock_scorer=3; ** IF SELF RESPONDENT **;


```{r}
clock_score_processing = function(sp,ROUND) {
  draw_str = paste('cg',ROUND,'dclkdraw',sep = '')
  speak_to_sp_str = paste('cg',ROUND,'speaktosp',sep = '')
  
  for (i in 1:nrow(sp)) {
    sp[i,'clock_test'] = sp[i,draw_str]
    
    if (sp[i,draw_str] %in% c(-1,-2) | is.na(sp[i,draw_str])) {
      sp[i,'clock_test'] = NA
    }
    
    else if (sp[i,draw_str] %in% c(-3,-4,-7)) {
      sp[i,'clock_test'] = 0
    }
    
    else if (sp[i, draw_str] == -9 & sp[i,speak_to_sp_str] == 1) {
      sp[i,'clock_test'] = 2
    }
    
    else if (sp[i, draw_str] == -9) {
      sp[i,'clock_test'] = NA
    }
  }
  return(sp)
}

round_1_cohort = clock_score_processing(round_1_cohort,1)
```

##Domain 3: memory (immediate word recall, delayed word recall)

```{r}

memory_processing = function(sp, ROUND) {
 
  imm_str = paste('cg',ROUND,'dwrdimmrc',sep = '')
  del_str = paste('cg',ROUND,'dwrddlyrc',sep = '')
  
  for (i in 1:2) {
    
    str = c(imm_str,del_str)[i]
    item_str = paste('mem_item_',i, sep = '') 
    
    sp[,item_str] = sp[,str]
    sp[sp[,str] %in% c(-1,-2), item_str] = NA
    sp[sp[,str] %in% c(-7,-3), item_str] = 0
    
  }
   
  sp$memory_domain = sp$mem_item_1 + sp$mem_item_2
  return(sp)
}

round_1_cohort = memory_processing(round_1_cohort, 1)
```

##Binarize domain scores + update dementia class with domain score info

```{r}

binary_processing = function(sp, ROUND) {
  
  sp$clock_binary = ifelse(sp$clock_test>1 & sp$clock_test<=5,0,1)
  sp$orientation_binary = ifelse(sp$orientation_domain>3 & sp$orientation_domain<=8,0,1)
  sp$memory_binary = ifelse(sp$memory_domain>3 & sp$memory_domain <= 20, 0,1)
  
  sp$domains_score = sp$clock_binary + sp$orientation_binary + sp$memory_binary
  
  for (i in 1:nrow(sp)) {
    if (is.na(sp[i,'dementia_class']) & sp[i,paste('cg',ROUND,'speaktosp',sep = '')] %in% c(-1,1)) {
      
      if (is.na(sp[i,'domains_score'])) {
        sp[i,'dementia_class'] = NA
      }
      
      else if (sp[i,'domains_score']>=2) {
        sp[i,'dementia_class'] = 2
      }
      
      else if (sp[i,'domains_score'] == 1) {
        sp[i,'dementia_class'] = 1
      }
      
      else if (sp[i,'domains_score'] == 0) {
        sp[i,'dementia_class'] = 0
      }
    }
  }
  
  return(sp)
}

round_1_cohort = binary_processing(round_1_cohort,1)

```

## STEADI derivation


```{r}
round_1_cohort = round_1_cohort %>%
  mutate(screen_1 =  case_when((hc1faleninyr == 2 | is.na(hc1faleninyr))  & 
                              (hc1worryfall == 2 | is.na(hc1worryfall))  & 
                              (ss1prbbalcrd ==2 | is.na(ss1prbbalcrd)) ~ 'Low risk',
                              hc1faleninyr == 1  |
                              hc1worryfall == 1  |
                              ss1prbbalcrd ==1 |
                                hc1multifall == 1 |
                                hc1faleninyr == 1 ~ 'Medium risk')) %>%
  mutate(screen_2 = case_when(screen_1 == 'Medium risk' & ba1ftdmreslt==2 ~ 'Potential high risk',
                              screen_1 == 'Medium risk' & ba1ftdmreslt==3 & (ba1rsn1ftstd==1 | ba1rsn2ftstd==1 | ba1rsn3ftstd==1) ~ 'Potential high risk',
                              screen_1 == 'Medium risk' & ((is.na(ba1ftdmreslt) & ba1dblftadm==3) | (is.na(ba1ftdmreslt) & ba1dblftadm==4))~ 'Potential high risk',
                              screen_1 == 'Medium risk' & ch12chstrslt==2 ~ 'Potential high risk',
                              screen_1 == 'Medium risk' & ch12chstrslt==3 &  (ch1chstntat1==1 | ch1chstntat2==1 | ch1chstntat3==1) ~ 'Potential high risk',
                              screen_1 == 'Medium risk' & ((is.na(ch12chstrslt) & ch1drchradm==3) | (is.na(ch12chstrslt) & ch1drchradm==4)) ~ 'Potential high risk')) %>%
  mutate(screen_3 = case_when(screen_2 == 'Potential high risk' & hc1multifall==1 ~ 'High risk',
                              screen_2 == 'Potential high risk' & hc1brokebon1==1 ~'High risk')) %>%
  mutate(STEADI_score = case_when(screen_1 == 'Low risk' ~ 'Low risk',
                                  screen_1 == 'Medium risk' & is.na(screen_2) & is.na(screen_3) ~ 'Medium risk',
                                  screen_1 == 'Medium risk' & screen_2 == 'Potential high risk' & is.na(screen_3) ~ 'Medium risk',
                                  screen_1 == 'Medium risk' & screen_3 == 'High risk' ~ 'High risk'))

```



#### Grouping 1 sarcopenia, obesity, sarcopenic obesity, neither
- Sarcopenia defined using grip strength cutoffs
- Obesity derived from BMI>=30 (variable name obesity)
- Sarcopenic obesity: both
- Neither
- appended as grouping_1_so_status

```{r}
round_1_cohort = round_1_cohort %>%
  mutate(grouping_1_so_status = case_when(sarcopenia == 1 & obesity == 0 ~ 'Sarcopenia',
                                          obesity == 1  &  sarcopenia == 0 ~ 'Obesity',
                                          sarcopenia == 1 & obesity == 1 ~ 'Sarcopenic Obesity',
                                          sarcopenia == 0 & obesity == 0 ~ 'Neither'))

#check
sum(is.na(round_1_cohort$grouping_1_so_status))
```

#### Grouping 2 sarcopenia, obesity, sarcopenic obesity, neither
- Sarcopenia defined using grip strength cutoffs
- Obesity derived from WC (variable name high_wc)
- Sarcopenic obesity: both
- Neither
- appended as grouping_2_so_status

```{r}
round_1_cohort = round_1_cohort %>%
  mutate(grouping_2_so_status = case_when(sarcopenia == 1 & high_wc == 0 ~ 'Sarcopenia',
                                          high_wc == 1 & sarcopenia == 0~ 'Obesity',
                                          sarcopenia == 1 & high_wc == 1 ~ 'Sarcopenic Obesity',
                                          sarcopenia == 0 & high_wc == 0 ~ 'Neither'))

#check
sum(is.na(round_1_cohort$grouping_2_so_status))
```

#### Export as csv
```{r}
#write.csv(round_1_cohort, 'processed_baseline.csv')
```