---
title: "NHATS_year_join"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, message = False, warning = False}
library(tidyverse)
library(readstata13)

round_1_cohort <- read.csv('processed_baseline.csv', header = T)
round_2_sp <- read.dta13('NHATS_Round_2_SP_File_v2.dta')
round_3_sp <- read.dta13('NHATS_Round_3_SP_File.dta')
round_4_sp <- read.dta13('NHATS_Round_4_SP_File.dta')
round_5_sp <- read.dta13('NHATS_Round_5B_SP_File.dta')
round_6_sp <- read.dta13('NHATS_Round_6_SP_File_V2.dta')
round_7_sp <- read.dta13('NHATS_Round_7_SP_File.dta')
round_8_sp <- read.dta13('NHATS_Round_8B_SP_File.dta')
```

##Round 2-8 preprocessing for AD8 dementia class

```{r}
ad8_processing = function(sp, ROUND) {
  for (i in 1:nrow(sp)) {
    
    ###inital codes for dementia_class
    #nursing home resident
    if (sp[i,'r1dresid'] == 7) {
      sp[i,'dementia_class'] = -9
    }                             
  
    if (sp[i,'r1dresid'] %in% c(6,8)) {
      sp[i,'dementia_class'] = -1
    }
    #default to NA
    else {
      sp[i,'dementia_class'] = NA
    }
    
    #default AD8 score and AD8 binary to NA
    sp[i,'ad8_score'] = NA
    sp[i,'ad8_binary'] = NA
    
    #Get AD8 score if respondent type is proxy and dementia class is NA
    if (sp[i,paste('is',ROUND,'resptype',sep = '')] == 2 & is.na(sp[i,'dementia_class'])) {
      
      sp[i,'ad8_score'] = 0
      
      for (j in 1:8) {
        
        item = paste('cp',ROUND,'chgthink',j,sep = '')
        recoded_item = paste('ad8_item_',j,sep = '')
        sp[i,recoded_item] = NA
        
        #Change reported by proxy (response 1) or dementia/alzheimer's reported by proxy (response 3)
        if (sp[i,item] %in% c(1,3)) {
          sp[i,recoded_item] = 1
        }
        
        #No change reported by proxy
        if (!is.na(sp[i,item]) & sp[i,item] == 2) {
          sp[i,recoded_item] = 0
        }
        
        sp[i,'ad8_score'] = sp[i,'ad8_score'] + sp[i,recoded_item]
      }
      
    }
    
    #fill inapplicable AD8 with -1 and AD8 items with NA
    else {
      for (j in 1:8) {
        recoded_item = paste('ad8_item_',j,sep = '')
        sp[i,recoded_item] = -1
        sp[i,paste('ad8_item_',j,sep = '')] = NA
      }
    }
    
    #if AD8 criterion met in previous round, then set AD8 score = 8 - not applicable for round 1
    if (ROUND != 1) {
      if (sp[i,paste('cp',ROUND,'dad8dem',sep = '')] == 1 & is.na(sp[i,'dementia_class'])) {
        sp[i, 'ad8_score'] = 8
      }
    }
    
    #create ad8_binary (0 - does not meet AD8 criterion, 1 - does meet AD8 criterion)
    if (sp[i,'ad8_score'] >= 2 & !is.na(sp[i,'ad8_score'])) {
      sp[i,'ad8_binary'] = 1
    }
    
    else if (sp[i, 'ad8_score'] %in% c(0,1) | 
             (all(is.na(subset(sp[i,],select = ad8_item_1:ad8_item_8))) & is.na(sp[i,'dementia_class']))) {
      
      sp[i,'ad8_binary'] = 0
    }
    
    #update dementia class
    
    if (is.na(sp[i, 'dementia_class']) & !is.na(sp[i,'ad8_binary']) & sp[i,'ad8_binary'] == 1) {
      sp[i,'dementia_class'] = 1
    }
    
    else if (is.na(sp[i, 'dementia_class']) & !is.na(sp[i,'ad8_binary']) & sp[i,'ad8_binary'] == 2 & sp[i,paste('cg',ROUND,'speaktosp',sep='')] ==2) {
      sp[i,'dementia_class'] = 3
    }
    
  }
  return(sp)
}  


round_2_processed = ad8_processing(round_2_sp,2)
round_3_processed = ad8_processing(round_3_sp,3)
round_4_processed = ad8_processing(round_4_sp,4)
round_5_processed = ad8_processing(round_5_sp,5)
round_6_processed = ad8_processing(round_6_sp,6)
round_7_processed = ad8_processing(round_7_sp,7)
round_8_processed = ad8_processing(round_8_sp,8)
```

##Round 2-8 preprocessing for domain 1: orientation (date recall, president naming)

cg*todaydat1: 
MONTH
-9 Missing
-7 RF
-1 Inapplicable
1 Yes
2 No/DK

cg*todaydat2:
DAY

cg*todaydat3:
YEAR

cg*todaydat4:
DAY OF WEEK

```{r}

orientation_domain_processing = function(sp, ROUND) {
  
  #dat recoding
  dat_string = paste('cg',ROUND,'todaydat', sep = '')
  
  if (ROUND != 4) {
    for (i in 1:4) {
      dat_str = paste(dat_string,i,sep = '')
      item_str = paste('date_item_',i, sep = '')
      sp[sp[,dat_str] %in% c(-7,2), item_str] = 0
      sp[sp[,dat_str] %in% c(-1,-9), item_str] = NA
      sp[sp[,dat_str] == 1, item_str] = 1
    }
  }
  
  else {
    for (i in c(1,2,3,5)) {
      dat_str = paste(dat_string,i,sep = '')
      item_str = paste('date_item_',i, sep = '')
      if (i == 5) {
        item_str = 'date_item_4'
      }
      sp[sp[,dat_str] %in% c(-7,2), item_str] = 0
      sp[sp[,dat_str] %in% c(-1,-9), item_str] = NA
      sp[sp[,dat_str] == 1, item_str] = 1
    }
  }
  
  sp$date_score = sp$date_item_1 + sp$date_item_2 + sp$date_item_3 + sp$date_item_4
  
  #all NAs in date_items and can't speak to SP (cg*speaktosp == 2) result in date_score == NA
  speaktosp = paste('cg',ROUND,'speaktosp',sep = '')
  all_na = is.na(sp$date_item_1) & is.na(sp$date_item_2) & is.na(sp$date_item_3) & is.na(sp$date_item_4)
  sp[sp[all_na, speaktosp] == 2,'date_score'] = NA
  
  
  pres_strings = c(paste('cg',ROUND,'presidna1',sep = ''),
                   paste('cg',ROUND,'presidna3',sep = ''),
                   paste('cg',ROUND,'vpname1',sep = ''),
                   paste('cg',ROUND,'vpname3',sep = ''))
  
  for (i in 1:4) {
    dat_str = pres_strings[i]
    item_str = paste('o_item_',i, sep = '')
    sp[sp[,dat_str] %in% c(-7,2), item_str] = 0
    sp[sp[,dat_str] %in% c(-1,-9), item_str] = NA
    sp[sp[,dat_str] == 1, item_str] = 1
  }
  
  sp$orientation_score = sp$o_item_1 + sp$o_item_2 + sp$o_item_3 + sp$o_item_4
  
  #all NAs in o_items and can't speak to SP (cg*speaktosp == 2) result in orientation_score == NA
  all_na_pres = is.na(sp$o_item_1) & is.na(sp$o_item_2) & is.na(sp$o_item_3) & is.na(sp$o_item_4)
  sp[sp[all_na_pres, speaktosp] == 2,'orientation_score'] = NA
  
  sp$orientation_domain = sp$date_score + sp$orientation_score
 
  return(sp)
}


round_2_processed = orientation_domain_processing(round_2_processed,2)
round_3_processed = orientation_domain_processing(round_3_processed,3)
round_4_processed = orientation_domain_processing(round_4_processed,4)
round_5_processed = orientation_domain_processing(round_5_processed,5)
round_6_processed = orientation_domain_processing(round_6_processed,6)
round_7_processed = orientation_domain_processing(round_7_processed,7)
round_8_processed = orientation_domain_processing(round_8_processed,8)

```


##Round 2-8 preprocessing for domain 2: executive function (clock score)

** 8) EXECUTIVE FUNCTION DOMAIN: CLOCK DRAWING SCORE **;
r2clock_scorer=cg2dclkdraw;
if cg2dclkdraw in (-2,-9) then r2clock_scorer=.;
if cg2dclkdraw in (-3,-4,-7) then r2clock_scorer=0;
** IMPUTE MEAN SCORE TO PERSONS MISSING A CLOCK **;
if cg2dclkdraw=-9 and cg2speaktosp=1 then r2clock_scorer=2; ** IF PROXY SAID CAN ASK SP **;
if cg2dclkdraw=-9 and cg2speaktosp=-1 then r2clock_scorer=3; ** IF SELF RESPONDENT **;


```{r}
clock_score_processing = function(sp,ROUND) {
  draw_str = paste('cg',ROUND,'dclkdraw',sep = '')
  speak_to_sp_str = paste('cg',ROUND,'speaktosp',sep = '')
  
  for (i in 1:nrow(sp)) {
    sp[i,'clock_test'] = sp[i,draw_str]
    
    if (sp[i,draw_str] %in% c(-1,-2) | is.na(sp[i,draw_str])) {
      sp[i,'clock_test'] = NA
    }
    
    else if (sp[i,draw_str] %in% c(-3,-4,-7)) {
      sp[i,'clock_test'] = 0
    }
    
    else if (sp[i, draw_str] == -9 & sp[i,speak_to_sp_str] == 1) {
      sp[i,'clock_test'] = 2
    }
    
    else if (sp[i, draw_str] == -9 & sp[i,speak_to_sp_str] == -1) {
      sp[i,'clock_test'] = 3
    }
    
    else if (sp[i, draw_str] == -9) {
      sp[i,'clock_test'] = NA
    }
  }
  return(sp)
}

round_2_processed = clock_score_processing(round_2_processed,2)
round_3_processed = clock_score_processing(round_3_processed,3)
round_4_processed = clock_score_processing(round_4_processed,4)
round_5_processed = clock_score_processing(round_5_processed,5)
round_6_processed = clock_score_processing(round_6_processed,6)
round_7_processed = clock_score_processing(round_7_processed,7)
round_8_processed = clock_score_processing(round_8_processed,8)
```


##Round 2-8 preprocessing for domain 3: memory (immediate word recall, delayed word recall)

```{r}

memory_processing = function(sp, ROUND) {
 
  imm_str = paste('cg',ROUND,'dwrdimmrc',sep = '')
  del_str = paste('cg',ROUND,'dwrddlyrc',sep = '')
  
  for (i in 1:2) {
    
    str = c(imm_str,del_str)[i]
    item_str = paste('mem_item_',i, sep = '') 
    
    sp[,item_str] = sp[,str]
    sp[sp[,str] %in% c(-1,-2), item_str] = NA
    sp[sp[,str] %in% c(-7,-3), item_str] = 0
    
  }
   
  sp$memory_domain = sp$mem_item_1 + sp$mem_item_2
  return(sp)
}

round_2_processed = memory_processing(round_2_processed,2)
round_3_processed = memory_processing(round_3_processed,3)
round_4_processed = memory_processing(round_4_processed,4)
round_5_processed = memory_processing(round_5_processed,5)
round_6_processed = memory_processing(round_6_processed,6)
round_7_processed = memory_processing(round_7_processed,7)
round_8_processed = memory_processing(round_8_processed,8)
```


##Binary categorization for three domains

Executive function domain (clock score) cutoffs:

1 < clock_test <= 5 then clock_binary = 0
0 <= clock_test <= 1 then clock_binary = 1

Orientation domain cutoffs:

3 < orientation_domain <= 8 then orientation_binary = 0
0 <= orientation_domain <= 3 then orientation_binary = 1

Memory domain cutoffs:

3 < memory_domain <= 20 then memory_binary = 0
0 <= memory_domain <= 3 then memory_binary = 1

Dementia class update rules:

Domains score >=2 then dementia_class = 2 (probable dementia)
Domains_score == 1 then dementia_class = 1 (possible dementia)
Domains_score = 0 then dementia_class = 0 (no dementia)


```{r}

binary_processing = function(sp, ROUND) {
  
  sp$clock_binary = ifelse(sp$clock_test>1 & sp$clock_test<=5,0,1)
  sp$orientation_binary = ifelse(sp$orientation_domain>3 & sp$orientation_domain<=8,0,1)
  sp$memory_binary = ifelse(sp$memory_domain>3 & sp$memory_domain <= 20, 0,1)
  
  sp$domains_score = sp$clock_binary + sp$orientation_binary + sp$memory_binary
  
  for (i in 1:nrow(sp)) {
    if (is.na(sp[i,'dementia_class']) & sp[i,paste('cg',ROUND,'speaktosp',sep = '')] %in% c(-1,1)) {
      
      if (is.na(sp[i,'domains_score'])) {
        sp[i,'dementia_class'] = NA
      }
      
      else if (sp[i,'domains_score']>=2) {
        sp[i,'dementia_class'] = 2
      }
      
      else if (sp[i,'domains_score'] == 1) {
        sp[i,'dementia_class'] = 1
      }
      
      else if (sp[i,'domains_score'] == 0) {
        sp[i,'dementia_class'] = 0
      }
    }
  }
  
  return(sp)
}

round_2_processed = binary_processing(round_2_processed,2)
round_3_processed = binary_processing(round_3_processed,3)
round_4_processed = binary_processing(round_4_processed,4)
round_5_processed = binary_processing(round_5_processed,5)
round_6_processed = binary_processing(round_6_processed,6)
round_7_processed = binary_processing(round_7_processed,7)
round_8_processed = binary_processing(round_8_processed,8)
```



##Join yearly data (rounds 2-8) to baseline preprocessed file

```{r}
round_1_scores = round_1_cohort %>%
  select(spid,
         grouping_1_so_status,
         grouping_2_so_status,
         sdoc_sarcopenia,
         sarcopenia,
         obesity,
         high_wc,
         gender = r1dgender,
         race,
         age_category,
         smoking_status,
         education,
         heart_disease,
         hypertension,
         arthritis,
         diabetes,
         lung_disease,
         stroke,
         cancer,
         ever_walk,
         vigor_actv,
         self_rated_memory = cg1ratememry,
         imm_recall = cg1dwrdimmrc,
         delayed_recall = cg1dwrddlyrc,
         clock_test,
         clock_binary,
         orientation_domain,
         orientation_binary,
         memory_domain,
         memory_binary,
         ad8_score,
         ad8_binary,
         dementia_class,
         STEADI_score)
                                       
round_2_scores = round_2_processed %>%
  select(spid,
         self_rated_memory = cg2ratememry,
         imm_recall = cg2dwrdimmrc,
         delayed_recall = cg2dwrddlyrc,
         clock_test,
         clock_binary,
         orientation_domain,
         orientation_binary,
         memory_domain,
         memory_binary,
         ad8_score,
         ad8_binary,
         dementia_class)

round_3_scores = round_3_processed %>%
  select(spid,
         self_rated_memory = cg3ratememry,
         imm_recall = cg3dwrdimmrc,
         delayed_recall = cg3dwrddlyrc,
         clock_test,
         clock_binary,
         orientation_domain,
         orientation_binary,
         memory_domain,
         memory_binary,
         ad8_score,
         ad8_binary,
         dementia_class)

round_4_scores = round_4_processed %>%
  select(spid,
         self_rated_memory = cg4ratememry,
         imm_recall = cg4dwrdimmrc,
         delayed_recall = cg4dwrddlyrc,
         clock_test,
         clock_binary,
         orientation_domain,
         orientation_binary,
         memory_domain,
         memory_binary,
         ad8_score,
         ad8_binary,
         dementia_class)

round_5_scores = round_5_processed %>%
  select(spid,
         self_rated_memory = cg5ratememry,
         imm_recall = cg5dwrdimmrc,
         delayed_recall = cg5dwrddlyrc,
         clock_test,
         clock_binary,
         orientation_domain,
         orientation_binary,
         memory_domain,
         memory_binary,
         ad8_score,
         ad8_binary,
         dementia_class)

round_6_scores = round_6_processed %>%
  select(spid,
         self_rated_memory = cg6ratememry,
         imm_recall = cg6dwrdimmrc,
         delayed_recall = cg6dwrddlyrc,
         clock_test,
         clock_binary,
         orientation_domain,
         orientation_binary,
         memory_domain,
         memory_binary,
         ad8_score,
         ad8_binary,
         dementia_class)

round_7_scores = round_7_processed %>%
  select(spid,
         self_rated_memory = cg7ratememry,
         imm_recall = cg7dwrdimmrc,
         delayed_recall = cg7dwrddlyrc,
         clock_test,
         clock_binary,
         orientation_domain,
         orientation_binary,
         memory_domain,
         memory_binary,
         ad8_score,
         ad8_binary,
         dementia_class)

round_8_scores = round_8_processed %>%
  select(spid,
         self_rated_memory = cg8ratememry,
         imm_recall = cg8dwrdimmrc,
         delayed_recall = cg8dwrddlyrc,
         clock_test,
         clock_binary,
         orientation_domain,
         orientation_binary,
         memory_domain,
         memory_binary,
         ad8_score,
         ad8_binary,
         dementia_class)

join_ = round_1_scores %>%
  left_join(round_2_scores, by = 'spid', suffix = c('_1','_2')) %>%
  left_join(round_3_scores, by = 'spid', suffix = c('_2','_3')) %>%
  left_join(round_4_scores, by = 'spid', suffix = c('_3','_4')) %>%
  left_join(round_5_scores, by = 'spid', suffix = c('_4','_5')) %>%
  left_join(round_6_scores, by = 'spid', suffix = c('_5','_6')) %>%
  left_join(round_7_scores, by = 'spid', suffix = c('_6','_7')) %>%
  left_join(round_8_scores, by = 'spid', suffix = c('_7','_8')) 

#write.csv(join_, 'joined_data.csv')
```